---
title: "Exercise scRNA-Seq"
author: "Anton Fornies"
date: "2025-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Matrix)
library(Seurat)
library(dplyr)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(tidyverse)
```

LOADING THE DATA

```{r}
matrix <- readMM("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Exp_data_UMIcounts.mtx")
dim(matrix)

barcodes <- read.csv("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Cells.csv")

features <- read.delim("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Genes.txt", header = F)

metadata <- read.csv("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Samples.csv")

```

MATRIX COUNTS

```{r}

rownames(matrix) <- features$V1
colnames(matrix) <- barcodes$cell_name

```

SEURAT OBJECT + METADATA

```{r}

seurat <- CreateSeuratObject(counts = matrix,
                             assay = "RNA",
                             names.delim = "_", 
                             names.field = 1:3)

seurat@meta.data$cell_name <- rownames(seurat@meta.data)

barcodes$nCount_RNA <- NULL
barcodes$nFeature_RNA <- NULL

seurat@meta.data <- seurat@meta.data %>%
  left_join(barcodes, by = "cell_name")

rownames(seurat@meta.data) <- seurat@meta.data$cell_name

```

QUALITY CONTROL

```{r}
# % mito genes
seurat[["percent.mt"]] <- PercentageFeatureSet(seurat, pattern = "^MT-")

# QC Graphs 
Idents(seurat) <- "orig.ident"
vln <- VlnPlot(seurat, features = "percent.mt") + theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\QC2.png", res = 550, width = 5500, height = 3000)
print(vln)
dev.off

scat <- FeatureScatter(seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + ggtitle("nCount_RNA vs nFeature_RNA")
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\QC1.png", res = 550, width = 5500, height = 3000)
print(scat)
dev.off

mean(seurat@meta.data$nCount_RNA)
mean(seurat@meta.data$nFeature_RNA)

# QC is already done but just to be sure
seurat <- subset(seurat,
                 subset = nFeature_RNA > 200 &
                          nFeature_RNA < 6000 &
                          nCount_RNA > 400 &
                          percent.mt < 15)
```

CELL CYCLE SCORE

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

```

DATA PRE-PROCESSING

```{r}
seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat)
seurat <- CellCycleScoring(seurat, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
seurat <- ScaleData(seurat, vars.to.regress = c("nCount_RNA","percent.mt","S.Score","G2M.Score"))
seurat <- RunPCA(seurat)

ElbowPlot(seurat, reduction = "pca")

seurat <- FindNeighbors(seurat, dims = 1:20, k.param = 20, compute.SNN = T, n.trees = 100)
seurat <- FindClusters(seurat, resolution = 0.8)

seurat <- RunUMAP(seurat, dims = 1:20)
```

Visualizations

```{r}
Idents(seurat) <- "seurat_clusters"

clusters <- DimPlot(seurat, reduction = "umap", label = TRUE)
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\clusters.png", res = 550, width = 5500, height = 3000)
print(clusters)
dev.off

orig_cell <- DimPlot(seurat, reduction = "umap", group.by = "cell_type")
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\orig_celltype.png", res = 550, width = 5500, height = 3000)
print(orig_cell)
dev.off

cc <- DimPlot(seurat, reduction = "umap", group.by = "Phase")
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\CC.png", res = 550, width = 5500, height = 3000)
print(cc)
dev.off

pat <- DimPlot(seurat, reduction = "umap", group.by = "sample")
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Patients.png", res = 550, width = 5500, height = 3000)
print(pat)
dev.off
```

Markers

```{r}
markers <- FindAllMarkers(seurat, test.use = "MAST", only.pos = T, latent.vars = c("nCount_RNA","percent.mt","S.Score","G2M.Score"))

markers <- markers %>%
  filter(p_val_adj < 0.05)
top_markers <- markers%>% group_by(cluster) %>% top_n(n = 400, wt = avg_log2FC)
```

Cell type annotation using scCATCH

```{r}
library(scCATCH)

data <- GetAssayData(seurat, layer = "data")

clusters <- seurat$seurat_clusters

sc <- createscCATCH(data = data, cluster = as.character(clusters))
sc <- findmarkergene(object = sc, species = "Human", marker = cellmatch, tissue = "Breast", cancer = "Breast Cancer")
result <- findcelltype(object = sc) # Clusters missing

```

scCATCH but markers found by MAST

```{r}
# With Markers reveal by FindAllMarkers
top_markers <- top_markers %>% ungroup()

sc@markergene <- top_markers

result <- findcelltype(object = sc)
```

With the predictions from both scCATCH and PanglaoDB

```{r}
pred_df <- data.frame(
  seurat_clusters = as.character(0:21),
  pred_celltype = c(
    "T Helper CD4+", "Malignant cancer cell", "T Cytotoxic CD8+ ", "Malignant cancer cell", "Macrophage", "Endothelial cell", "T Helper CD4+", "Malignant cancer cell", "Malignant cancer cell", "Malignant cancer cell", "Dendritic cell", "B cell", "Natural Killers", "Fibroblast", "Malignant cancer cell", "Malignant cancer cell", "Malignant cancer cell", "Malignant cancer cell", "Malignant cancer cell", "T Helper CD4+", "Malignant cancer cell", "T Helper CD4+"),
  stringsAsFactors = FALSE
)

# Convertir seurat_clusters en character si aún no lo es
seurat@meta.data$seurat_clusters <- as.character(seurat@meta.data$seurat_clusters)

# Merge con metadata
seurat@meta.data <- left_join(
  seurat@meta.data,
  pred_df,
  by = "seurat_clusters"
)
rownames(seurat@meta.data) <- seurat@meta.data$cell_name

celltypes <- DimPlot(
  seurat,
  group.by = "pred_celltype"
) +
  ggtitle("UMAP of Predicted Cell Types") +
  xlab("UMAP 1") +
  ylab("UMAP 2") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10), 
    legend.position = "bottom"
  )
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Celltypes.png", res = 550, width = 5000, height = 3800)
print(celltypes)
dev.off

table(seurat@meta.data$pred_celltype)
table(barcodes$cell_type)

```

Cell type annotation using SingleR

```{r}

library(SingleR)
library(celldex)  
expr <- GetAssayData(seurat, slot = "data")
meta <- seurat@meta.data

ref <- celldex::DatabaseImmuneCellExpressionData()

pred <- SingleR(test = expr, ref = ref, labels = ref$label.main)

seurat$SingleR_label <- pred$labels[match(rownames(seurat@meta.data), rownames(pred))]

# Monocytes confounded with cancer cells
seurat@meta.data$SingleR_label[seurat@meta.data$SingleR_label == "Malignant cells"] <- "Monocytes"

DimPlot(seurat, group.by = "SingleR_label",split.by = "SingleR_label")

# Evaluate perform of singleR
p <- plotScoreHeatmap(pred)
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\heat_singleR.png", res = 550, width = 4500, height = 3500)
print(p)
dev.off

plotDeltaDistribution(pred)

# Cluster correspondence
tab <- table(Assigned=pred$labels, Cluster=seurat$seurat_clusters)
tab

library(pheatmap)
pheatmap(log(tab+10))

singleR <- DimPlot(seurat, group.by = "SingleR_label")
png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\singleR2.png", res = 550, width = 5500, height = 3000)
print(singleR)
dev.off

table(seurat@meta.data$pred_celltype)
table(barcodes$cell_type)
table(seurat@meta.data$SingleR_label)

```

Comparing cancer subtypes: "TNBC" vs "ER+/PR+/HER2-"
Pre and On samples: 2 time points

```{r}
# Metadata Cancer Subtype
metadata$sample <- as.character(metadata$sample)
seurat@meta.data$sample <- as.character(seurat@meta.data$sample)

bc_map <- setNames(metadata$mutation_hormonal_subtype, metadata$sample)

seurat@meta.data$Cancer_Subtype <- bc_map[seurat@meta.data$sample]
table(seurat@meta.data$Cancer_Subtype)

Idents(seurat) <- "Cancer_Subtype"

rownames(seurat@assays[["RNA"]]@layers[["counts"]]) <- rownames(seurat)
counts_matrix <- seurat@assays[["RNA"]]@layers[["counts"]]
gene_counts <- Matrix::rowSums(counts_matrix > 0)
genes_to_keep <- names(gene_counts[gene_counts >= 400])
seurat <- subset(seurat, features = genes_to_keep)

seurat_On <- subset(seurat, subset = grepl("On$", orig.ident))
seurat_Pre <- subset(seurat, subset = grepl("Pre$", orig.ident))

DEGs_Pre <- FindMarkers(seurat_Pre, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-", test.use = "MAST", latent.vars = c("nCount_RNA","percent.mt","S.Score","G2M.Score"))
DEGs_Pre$Gene <- rownames(DEGs_Pre)
write.xlsx(DEGs_Pre, "C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\DEGs_pre.xlsx")

library(EnhancedVolcano)
library(ggrepel)
vl1 <- EnhancedVolcano(DEGs_Pre,
  lab = DEGs_Pre$Gene,
  x = 'avg_log2FC',
  y = 'p_val',
  pCutoff = 0.05,
  FCcutoff = 0.5,
  xlim = c(-3, 3),
  pointSize = 2.5,
  labSize = 3,  
  colAlpha = 0.75,
  legendPosition = 'right',
  title = 'Comparing Breast cancer subtypes',
  subtitle = 'TNBC vs ER+/PR+/HER2-',
  boxedLabels = TRUE,  
  drawConnectors = TRUE,
  widthConnectors = 0.3
) +
  theme(legend.position = "bottom", 
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)
  )


png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Volcano_PRE.png", res = 550, width = 5500, height = 4200)
print(vl1)
dev.off

###############################################################################

DEGs_On <- FindMarkers(seurat_On, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-", test.use = "MAST", latent.vars = c("nCount_RNA","percent.mt","S.Score","G2M.Score"))
DEGs_On$Gene <- rownames(DEGs_On)
write.xlsx(DEGs_On, "C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\DEGs_on.xlsx")

vl2 <- EnhancedVolcano(DEGs_On,
                lab = DEGs_On$Gene,
               x = 'avg_log2FC',
  y = 'p_val',
  pCutoff = 0.05,
  FCcutoff = 0.5,
  xlim = c(-3, 3),
  pointSize = 2.5,
  labSize = 3,  
  colAlpha = 0.75,
  legendPosition = 'right',
  title = 'Comparing Breast cancer subtypes',
  subtitle = 'TNBC vs ER+/PR+/HER2-',
  boxedLabels = TRUE,  
  drawConnectors = TRUE,
  widthConnectors = 0.3
) +
  theme(legend.position = "bottom", 
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)
  )

png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\Volcano_ON.png", res = 550, width = 5500, height = 4200)
print(vl2)
dev.off

#####################################################################3

top_pre <- DEGs_Pre %>%
  filter(p_val < 0.05 & abs(avg_log2FC) > 0.5) %>%
  mutate(abs_log2FC = abs(avg_log2FC)) %>%
  arrange(desc(abs_log2FC))

top_on <- DEGs_On %>%
  filter(p_val < 0.05 & abs(avg_log2FC) > 0.5) %>%
  mutate(abs_log2FC = abs(avg_log2FC)) %>%
  arrange(desc(abs_log2FC))

both <- intersect(top_pre, top_on)

```


For each cell type

Pre-treatment

```{r}
Idents(seurat_Pre) <- "Cancer_Subtype"

# B cell
subset_bcell <- subset(seurat_Pre, subset = pred_celltype == "B cell")
Idents(subset_bcell) <- "Cancer_Subtype"
degs_bcell <- FindMarkers(subset_bcell,
                          ident.1 = "TNBC",
                          ident.2 = "ER+/PR+/HER2-",
                          test.use = "MAST",
                          latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_bcell$gene <- rownames(degs_bcell)
degs_bcell$celltype <- "B cell"

# Dendritic cell
subset_dendritic <- subset(seurat_Pre, subset = pred_celltype == "Dendritic cell")
Idents(subset_dendritic) <- "Cancer_Subtype"
degs_dendritic <- FindMarkers(subset_dendritic,
                              ident.1 = "TNBC",
                              ident.2 = "ER+/PR+/HER2-",
                              test.use = "MAST",
                              latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_dendritic$gene <- rownames(degs_dendritic)
degs_dendritic$celltype <- "Dendritic cell"

# Endothelial cell
subset_endothelial <- subset(seurat_Pre, subset = pred_celltype == "Endothelial cell")
Idents(subset_endothelial) <- "Cancer_Subtype"
degs_endothelial <- FindMarkers(subset_endothelial,
                                ident.1 = "TNBC",
                                ident.2 = "ER+/PR+/HER2-",
                                test.use = "MAST",
                                latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_endothelial$gene <- rownames(degs_endothelial)
degs_endothelial$celltype <- "Endothelial cell"

# Fibroblast
subset_fibroblast <- subset(seurat_Pre, subset = pred_celltype == "Fibroblast")
Idents(subset_fibroblast) <- "Cancer_Subtype"
degs_fibroblast <- FindMarkers(subset_fibroblast,
                               ident.1 = "TNBC",
                               ident.2 = "ER+/PR+/HER2-",
                               test.use = "MAST",
                               latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_fibroblast$gene <- rownames(degs_fibroblast)
degs_fibroblast$celltype <- "Fibroblast"

# Macrophage
subset_macrophage <- subset(seurat_Pre, subset = pred_celltype == "Macrophage")
Idents(subset_macrophage) <- "Cancer_Subtype"
degs_macrophage <- FindMarkers(subset_macrophage,
                               ident.1 = "TNBC",
                               ident.2 = "ER+/PR+/HER2-",
                               test.use = "MAST",
                               latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_macrophage$gene <- rownames(degs_macrophage)
degs_macrophage$celltype <- "Macrophage"

# Malignant cancer cell
subset_malignant <- subset(seurat_Pre, subset = pred_celltype == "Malignant cancer cell")
Idents(subset_malignant) <- "Cancer_Subtype"
degs_malignant <- FindMarkers(subset_malignant,
                               ident.1 = "TNBC",
                               ident.2 = "ER+/PR+/HER2-",
                               test.use = "MAST",
                               latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_malignant$gene <- rownames(degs_malignant)
degs_malignant$celltype <- "Malignant cancer cell"

# Natural Killers
subset_nk <- subset(seurat_Pre, subset = pred_celltype == "Natural Killers")
Idents(subset_nk) <- "Cancer_Subtype"
degs_nk <- FindMarkers(subset_nk,
                       ident.1 = "TNBC",
                       ident.2 = "ER+/PR+/HER2-",
                       test.use = "MAST",
                       latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_nk$gene <- rownames(degs_nk)
degs_nk$celltype <- "Natural Killers"

# T Cytotoxic CD8+
subset_cd8 <- subset(seurat_Pre, subset = pred_celltype == "T Cytotoxic CD8+ ")
Idents(subset_cd8) <- "Cancer_Subtype"
degs_cd8 <- FindMarkers(subset_cd8,
                        ident.1 = "TNBC",
                        ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_cd8$gene <- rownames(degs_cd8)
degs_cd8$celltype <- "T Cytotoxic CD8+"

# T Helper CD4+
subset_cd4 <- subset(seurat_Pre, subset = pred_celltype == "T Helper CD4+")
Idents(subset_cd4) <- "Cancer_Subtype"
degs_cd4 <- FindMarkers(subset_cd4,
                        ident.1 = "TNBC",
                        ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_cd4$gene <- rownames(degs_cd4)
degs_cd4$celltype <- "T Helper CD4+"

# Bind all
DEGs_celltype_Pre <- bind_rows(
  degs_bcell,
  degs_dendritic,
  degs_endothelial,
  degs_fibroblast,
  degs_macrophage,
  degs_malignant,
  degs_nk,
  degs_cd8,
  degs_cd4
)

write.xlsx(DEGs_celltype_Pre, "C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\DEGs_celltype_pre.xlsx")

```

On-treatment

```{r}
Idents(seurat_On) <- "Cancer_Subtype"

# B cell
subset_b <- subset(seurat_On, subset = pred_celltype == "B cell")
Idents(subset_b) <- "Cancer_Subtype"
degs_b <- FindMarkers(subset_b, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                      test.use = "MAST",
                      latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_b$gene <- rownames(degs_b)
degs_b$celltype <- "B cell"

# Dendritic cell
subset_dc <- subset(seurat_On, subset = pred_celltype == "Dendritic cell")
Idents(subset_dc) <- "Cancer_Subtype"
degs_dc <- FindMarkers(subset_dc, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                       test.use = "MAST",
                       latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_dc$gene <- rownames(degs_dc)
degs_dc$celltype <- "Dendritic cell"

# Endothelial cell
subset_endo <- subset(seurat_On, subset = pred_celltype == "Endothelial cell")
Idents(subset_endo) <- "Cancer_Subtype"
degs_endo <- FindMarkers(subset_endo, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                         test.use = "MAST",
                         latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_endo$gene <- rownames(degs_endo)
degs_endo$celltype <- "Endothelial cell"

# Fibroblast
subset_fib <- subset(seurat_On, subset = pred_celltype == "Fibroblast")
Idents(subset_fib) <- "Cancer_Subtype"
degs_fib <- FindMarkers(subset_fib, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_fib$gene <- rownames(degs_fib)
degs_fib$celltype <- "Fibroblast"

# Macrophage
subset_mac <- subset(seurat_On, subset = pred_celltype == "Macrophage")
Idents(subset_mac) <- "Cancer_Subtype"
degs_mac <- FindMarkers(subset_mac, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_mac$gene <- rownames(degs_mac)
degs_mac$celltype <- "Macrophage"

# Malignant cancer cell
subset_mal <- subset(seurat_On, subset = pred_celltype == "Malignant cancer cell")
Idents(subset_mal) <- "Cancer_Subtype"
degs_mal <- FindMarkers(subset_mal, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_mal$gene <- rownames(degs_mal)
degs_mal$celltype <- "Malignant cancer cell"

# Natural Killers
subset_nk <- subset(seurat_On, subset = pred_celltype == "Natural Killers")
Idents(subset_nk) <- "Cancer_Subtype"
degs_nk <- FindMarkers(subset_nk, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                       test.use = "MAST",
                       latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_nk$gene <- rownames(degs_nk)
degs_nk$celltype <- "Natural Killers"

# T Cytotoxic CD8+
subset_cd8 <- subset(seurat_On, subset = pred_celltype == "T Cytotoxic CD8+ ")
Idents(subset_cd8) <- "Cancer_Subtype"
degs_cd8 <- FindMarkers(subset_cd8, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_cd8$gene <- rownames(degs_cd8)
degs_cd8$celltype <- "T Cytotoxic CD8+"

# T Helper CD4+
subset_cd4 <- subset(seurat_On, subset = pred_celltype == "T Helper CD4+")
Idents(subset_cd4) <- "Cancer_Subtype"
degs_cd4 <- FindMarkers(subset_cd4, ident.1 = "TNBC", ident.2 = "ER+/PR+/HER2-",
                        test.use = "MAST",
                        latent.vars = c("nCount_RNA", "percent.mt", "S.Score", "G2M.Score"))
degs_cd4$gene <- rownames(degs_cd4)
degs_cd4$celltype <- "T helper CD4+"

DEGs_celltype_On <- bind_rows(
  degs_b,
  degs_dc,
  degs_endo,
  degs_fib,
  degs_mac,
  degs_mal,
  degs_nk,
  degs_cd8,
  degs_cd4
)

write.xlsx(DEGs_celltype_On, "C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\DEGs_celltype_On.xlsx")


```

HeatMaps per cell type

```{r}
library(ComplexHeatmap)

top_pre_10 <- top_pre %>% slice_head(n = 10)

DEGs_celltype_Pre_TOP <- DEGs_celltype_Pre %>%
  filter(gene %in% top_pre_10$Gene)

heatmap_data_pre <- DEGs_celltype_Pre_TOP %>%
  dplyr::select(gene, celltype, avg_log2FC) %>%
  pivot_wider(names_from = celltype, values_from = avg_log2FC) %>%
  column_to_rownames("gene") %>%
  as.matrix()

heatmap_data_pre[is.na(heatmap_data_pre)] <- 0

heatmap_pre <- Heatmap(
  heatmap_data_pre,
  col = colorRamp2(c(-3.5, 0, 3.5), c("steelblue", "white", "tomato")),
  name = "log2FC",
  rect_gp = gpar(col = "grey", lwd = 2),
  row_names_gp = gpar(fontsize = 11),
  row_names_side = "left",
  column_names_gp = gpar(fontsize = 12),
  column_names_rot = 90,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  clustering_method_columns = "ward.D2",
  clustering_distance_columns = "euclidean",
  column_title = "Specific cell types",
  column_title_gp = gpar(fontsize = 10)
)

order <- rownames(heatmap_data_pre)

# Bulk regulation
bulk_sign <- top_pre_10 %>%
  mutate(`Bulk dysregulation` = ifelse(avg_log2FC > 0, "Upregulation", "Downregulation")) %>%
  dplyr::select(`Bulk dysregulation`) %>%
  mutate(`Bulk dysregulation` = factor(`Bulk dysregulation`, levels = c("Downregulation", "Upregulation"))) %>%
  as.matrix()

bulk_sign <- bulk_sign[order, , drop = FALSE]

regulation_colors <- c("Downregulation" = "steelblue", "Upregulation" = "tomato")

heatmap_pre_bin <- Heatmap(
  bulk_sign,
  width = unit(0.8, "cm"),
  col = regulation_colors,
  name = "Regulation",
  rect_gp = gpar(col = "grey", lwd = 2),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  row_names_gp = gpar(fontsize = 11),
  column_names_gp = gpar(fontsize = 12),
  column_title = "Bulk regulation",
  column_title_gp = gpar(fontsize = 10)
)
draw(heatmap_pre + heatmap_pre_bin, heatmap_legend_side = "right")

png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\heatmap_pre.png", res = 550, width = 6000, height = 4000)
draw(heatmap_pre + heatmap_pre_bin, heatmap_legend_side = "right")
dev.off
```


```{r}
######################################################################33

top_on_10 <- top_on %>% slice_head(n = 10)

DEGs_celltype_On_TOP <- DEGs_celltype_On %>%
  filter(gene %in% top_on_10$Gene)

heatmap_data_on <- DEGs_celltype_On_TOP %>%
  dplyr::select(gene, celltype, avg_log2FC) %>%
  pivot_wider(names_from = celltype, values_from = avg_log2FC) %>%
  column_to_rownames("gene") %>%
  as.matrix()

heatmap_data_on[is.na(heatmap_data_on)] <- 0

heatmap_on <- Heatmap(
  heatmap_data_on,
  col = colorRamp2(c(-3.5, 0, 3.5), c("steelblue", "white", "tomato")),
  name = "log2FC",
  rect_gp = gpar(col = "grey", lwd = 2),
  row_names_gp = gpar(fontsize = 11),
  row_names_side = "left",
  column_names_gp = gpar(fontsize = 12),
  column_names_rot = 90,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  clustering_method_columns = "ward.D2",
  clustering_distance_columns = "euclidean",
  column_title = "Specific cell types",
  column_title_gp = gpar(fontsize = 10)
)

order <- rownames(heatmap_data_on)

# Bulk regulation
bulk_sign <- top_on_10 %>%
  mutate(`Bulk dysregulation` = ifelse(avg_log2FC > 0, "Upregulation", "Downregulation")) %>%
  dplyr::select(`Bulk dysregulation`) %>%
  mutate(`Bulk dysregulation` = factor(`Bulk dysregulation`, levels = c("Downregulation", "Upregulation"))) %>%
  as.matrix()

bulk_sign <- bulk_sign[order, , drop = FALSE]

regulation_colors <- c("Downregulation" = "steelblue", "Upregulation" = "tomato")

heatmap_on_bin <- Heatmap(
  bulk_sign,
  width = unit(0.8, "cm"),
  col = regulation_colors,
  name = "Regulation",
  rect_gp = gpar(col = "grey", lwd = 2),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  row_names_gp = gpar(fontsize = 11),
  column_names_gp = gpar(fontsize = 12),
  column_title = "Bulk regulation",
  column_title_gp = gpar(fontsize = 10)
)
draw(heatmap_on + heatmap_on_bin, heatmap_legend_side = "right")


png("C:\\Users\\aforn\\OneDrive\\Documents\\SingleCell\\heatmap_on.png", res = 550, width = 6000, height = 4000)
draw(heatmap_on + heatmap_on_bin, heatmap_legend_side = "bottom")
dev.off

```

